---
title: "Generalized nonparametric regression in reproducing kernel Hilbert spaces"
output: github_document
---

## Download and source the functions

This repository contains fast ```C++``` implementations with an ```R``` interface of the reproducing kernel Hilbert space estimators of Kalogridis (2026).

The computation is done with **Iteratively Reweighted Least-Squares** and the penalty parameter is selected with **robust Generalized Cross Validation**.

Here are detailed installation instructions:

1. First, download all the files in your ```R``` working directory. This is at:

```{r echo=TRUE, message=FALSE, results = "hide"}
getwd()
```
2. Load the ```R``` function ```rkhs.R``` (Quantile and Least-squares RKHS estimators).

```{r message = FALSE}
source("rkhs.R")    # Quantile and LS estimators
```

3. Be sure to have installed and loaded the ```R```-packages ```Rcpp``` and ```RcppArmadillo```:

```{r, eval = FALSE}
# install.packages(c("Rcpp", "RcppArmadillo"))
library(Rcpp); library(RcppArmadillo)
```

4. The ```R```-functions will source the ```rkhs_quan.cpp``` file containing the ```C++``` implementations; no ```C``` knowledge is required.

## Example 1: One-dimensional data

```{r message=FALSE, warning = FALSE, cache = FALSE}
set.seed(10)

n    <- 200 # 50 for smaller samples
beta <- 0.75 # one of c(0.5, 0.75, 1.0) 
p <- 1.2

K <- 50
k <- 1:K

coeff <- k^(-(2 * beta + p))

f0 <- function(X) {
  X <- as.matrix(X)
  out <- rep(0, nrow(X))
  for (j in 1:K) {
    out <- out + coeff[j] * cos(2 * pi * j * X[,1])
  }
  return(out)}

X <- matrix(runif(n), n, 1)

# Targets
Xg <-  seq(0, 1, length = 300)
f0g <- f0(Xg)

y <- f0(X) + rnorm(n)

fit_ls <- fit_rkhs(
  X, y, loss = "ls",
  kernel = "matern")

fit_q <- fit_rkhs(
  X, y, loss = "quantile", tau = 0.5,
  kernel = "matern")

f_ls <- predict_rkhs(fit_ls, X, Xg)
f_q  <- predict_rkhs(fit_q,  X, Xg)

```

```{r message=FALSE, warning=FALSE, fig.width=8, fig.height=5, dpi=120}
library(ggplot2)

df <- data.frame(
  t = rep(Xg, 3),
  value = c(f0g, f_q, f_ls),
  method = factor(rep(c("True","LAD", "LS"), each=length(Xg)),
                  levels = c("True", "LAD", "LS")))

colors <- c("True" = "black", "LAD" = "red", "LS" = "blue")
line_types <- c("True" = "solid", "LAD" = "longdash", "LS" = "twodash")

ggplot(df, aes(x=t, y=value, color=method, linetype=method)) +
  geom_line(size=1.2) +
  theme_minimal(base_size = 16) +
  labs(x="x", y="", title="") + coord_cartesian(ylim=c(-1.2,1.2)) + 
  scale_color_manual(values = colors) +
  scale_linetype_manual(values = line_types) + 
  theme(legend.title = element_blank(),
        legend.position = "bottom",
        plot.title = element_text(hjust = 0.5))
```

If the measurement errors follow a light-tailed distribution, the estimators perform comparably. 

But for heavy tailed errors the situation changes dramatically in favour of the robust quantile estimator:

```{r message=FALSE, warning = FALSE, cache=FALSE}
set.seed(10)

y <- f0(X) + rt(n, df = 1)

fit_ls <- fit_rkhs(
  X, y, loss = "ls",
  kernel = "matern")

fit_q <- fit_rkhs(
  X, y, loss = "quantile", tau = 0.5,
  kernel = "matern")

f_ls <- predict_rkhs(fit_ls, X, Xg)
f_q  <- predict_rkhs(fit_q,  X, Xg)

df <- data.frame(
  t = rep(Xg, 3),
  value = c(f0g, f_q, f_ls),
  method = factor(rep(c("True","LAD", "LS"), each=length(Xg)),
                  levels = c("True", "LAD", "LS")))

colors <- c("True" = "black", "LAD" = "red", "LS" = "blue")
line_types <- c("True" = "solid", "LAD" = "longdash", "LS" = "twodash")

ggplot(df, aes(x=t, y=value, color=method, linetype=method)) +
  geom_line(size=1.2) +
  theme_minimal(base_size = 16) +
  labs(x="x", y="", title="") + coord_cartesian(ylim=c(-1.2,1.2)) + 
  scale_color_manual(values = colors) +
  scale_linetype_manual(values = line_types) + 
  theme(legend.title = element_blank(),
        legend.position = "bottom",
        plot.title = element_text(hjust = 0.5))

```



## Example 2: Two-dimensional data



## Notes

- See the ```R```-functions for complete documentation of the settings/options.
- The simulation script reproduces the results from Section 6 in Kalogridis (2026).

Contact me at ioannis.kalogridis@glasgow.ac.uk for any issues, questions or suggestions.

## References
Kalogridis, I. (2026). Generalized nonparametric regression in reproducing kernel Hilbert spaces: Consistency and rates of convergence, under review.
