---
title: "Generalized nonparametric regression in reproducing kernel Hilbert spaces"
output: github_document
---

This repository contains fast ```C++``` implementations with an ```R``` interface of the reproducing kernel Hilbert space estimators of Kalogridis (2026).

The computation is done with **Iteratively Reweighted Least-Squares** and the penalty parameter is selected with **robust Generalized Cross Validation**.

Here are detailed installation instructions:

1. First, download all the files in your ```R``` working directory. This is at:

```{r echo=TRUE, message=FALSE, results = "hide"}
getwd()
```
2. Load the ```R``` function ```rkhs.R``` (Quantile and Least-squares RKHS estimators).

```{r message = FALSE}
source("rkhs.R")    # Quantile and LS estimators
```

3. Be sure to have installed and loaded the ```R```-packages ```fda```, ```Rcpp``` and ```RcppArmadillo```:

```{r, eval = FALSE}
# install.packages(c("fda", "Rcpp", "RcppArmadillo"))
library(fda);library(Rcpp);library(RcppArmadillo)
```

4. The ```R```-functions will source the ```rkhs_quan.cpp``` file containing the ```C++``` implementations; no ```C``` knowledge is required.

## Notes

- Please see the ```R```-functions for complete documentation of the settings/options.
- The simulation script reproduces the results from Section 6 in Kalogridis (2026).

Get in contact with me at ioannis.kalogridis@glasgow.ac.uk for any issues, questions or suggestions.

## References
Kalogridis, I. (2026). Generalized nonparametric regression in reproducing kernel Hilbert spaces: Consistency and rates of convergence, under review.
